---
title: "simpleRegress"
author: "David Barton"
date: "Fall 2025"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## make a simple regression model

```{r}
bears <- read.csv(file="BEARS.csv",header = TRUE)
#adjust factors to be factors
bears$SEX<- as.factor(bears$SEX)
bears$MONTH<- as.factor(bears$MONTH)
```

## Model WEIGHT ~ NECK  this is an example  

```{r}
simp <- lm(WEIGHT~NECK,data=bears)
plot(WEIGHT~NECK,data=bears)
abline(simp)
summary.lm(simp)
```


## Compare all models that predict WEIGHT other than NECK. Then order them from worst to best by its value of Adjusted-Rsquared (bigger is better), or by RSE (residual standard error) where smaller is better. 

```{r}
simp1 <- lm(WEIGHT~CHEST,data=bears)
plot(WEIGHT~CHEST,data=bears)
abline(simp1)
summary.lm(simp)
```
The predicted weight overall of the bears based on to the chest. Adjusted R-squared value of 0.9262, and the Residual standard error of 33.08

```{r}
simp1 <- lm(WEIGHT~LENGTH,data=bears)
plot(WEIGHT~LENGTH,data=bears)
abline(simp1)
summary.lm(simp)
```

The prediction of the overall weight has adjusted R-square vaule of 0.87 and a Residual Standard error of 43.91

```{r}
simp3 <- lm(WEIGHT~HEADWTH,data=bears)
plot(WEIGHT~LENGTH,data=bears)
abline(simp3)
summary.lm(simp)
```

The overall weight of the bear based on the head width has an adjusted R-square value of 0.87 and a REsidual standard error vaule of  76.41

```{r}
simp4 <- lm(WEIGHT~HEADLEN,data=bears)
plot(WEIGHT~HEADLEN,data=bears)
abline(simp4)
summary.lm(simp)
```

The overall weight of the bear based on the head width has an adjusted R-square value of 0.69 and a REsidual standard error value of 67.81

```{r}
simp5 <- lm(WEIGHT~AGE,data=bears)
plot(WEIGHT~AGE,data=bears)
abline(simp5)
summary.lm(simp)
```

The overall weight of the bear based on the age has an adjusted R-square value of 0.5526 and a Residual standard error value of 81.47

